\PassOptionsToClass{BCOR=5mm, twoside}{tudapub}

\documentclass[english, notodo]{fdsummary}

\subtitle{Bachelor of Science (Part 2)}
\setcounter{tocdepth}{1}

\newcommand{\bspw}{bspw.\xspace}
\newcommand{\bzgl}{bzgl.\xspace}
\newcommand{\bzw}{bzw.\xspace}
\newcommand{\ca}{ca.\xspace}
\newcommand{\gdw}{gdw.\xspace}
\newcommand{\ggf}{ggf.\xspace}
\newcommand{\iA}{i.\,A.\xspace}
\newcommand{\mglw}{mglw.\xspace}
\newcommand{\vs}{vs.\xspace}
\newcommand{\zB}{z.\,B.\xspace}
\renewcommand{\dh}{d.\,h.\xspace}
\newcommand{\ggT}{\ensuremath{\textrm{ggT}}}

\newcommand{\lstbasepath}{unset}

\begin{document}
	\selectlanguage{USenglish}
	\maketitle
	\tableofcontents
	
	% 5. Semester: CSS, Mathe 3, BS
	\cleardoublepage
	\selectlanguage{ngerman}
	\part{Computersystemsicherheit}
	\graphicspath{{./cs/mandatory/fs3/css-fischlin/}}
	\renewcommand{\lstbasepath}{./cs/mandatory/fs3/css-fischlin}
	\include{cs/mandatory/fs3/css-fischlin/content}
	
	\cleardoublepage
	\selectlanguage{ngerman}
	\part{Mathematik 3 (Numerik)}
	\graphicspath{{./cs/mandatory/fs4/mathe3-numerik/}}
	\renewcommand{\lstbasepath}{./cs/mandatory/fs4/mathe3-numerik}
	\include{cs/mandatory/fs4/mathe3-numerik/content}
	\cleardoublepage
	\selectlanguage{ngerman}
	\part{Mathematik 3 (Stochastik/Statistik)}
	\graphicspath{{./cs/mandatory/fs4/mathe3-stochastik/}}
	\renewcommand{\lstbasepath}{./cs/mandatory/fs4/mathe3-stochastik}
	\include{cs/mandatory/fs4/mathe3-stochastik/content}
	
	\cleardoublepage
	\selectlanguage{USenglish}
	\part{Operating Systems}
	\graphicspath{{./cs/mandatory/fs5/os/}}
	\renewcommand{\lstbasepath}{./cs/mandatory/fs5/os}
	\include{cs/mandatory/fs5/os/content}
	
	% 6. Semester: SML, FMSe
	\cleardoublepage
	\selectlanguage{USenglish}
	\part{Statistical Machine Learning}
	\graphicspath{{./cs/elective/vc/statml/}}
	\renewcommand{\lstbasepath}{./cs/elective/vc/statml}
	\include{cs/elective/vc/statml/content}
	
	\cleardoublepage
	\selectlanguage{USenglish}
	\part{Formale Methoden im Softwareentwurf}
	\graphicspath{{./cs/mandatory/fs4/fmse/}}
	\renewcommand{\lstbasepath}{./cs/mandatory/fs4/fmse}
	\include{cs/mandatory/fs4/fmse/content}
	
	% 7. Semester: GdR, VC
	\cleardoublepage
	\selectlanguage{ngerman}
	\part{Grundlagen der Robotik}
	\graphicspath{{./cs/elective/ce/gdr/}}
	\renewcommand{\lstbasepath}{./cs/elective/ce/gdr}
	\include{cs/elective/ce/gdr/content}
	
	\cleardoublepage
	\selectlanguage{ngerman}
	\part{Visual Computing}
	\graphicspath{{./cs/mandatory/fs5/vc/}}
	\renewcommand{\lstbasepath}{./cs/mandatory/fs5/vc}
	\include{cs/mandatory/fs5/vc/content}
\end{document}
