\PassOptionsToClass{BCOR=5mm, twoside}{tudapub}

\documentclass[english, notodo]{fdsummary}

\subtitle{Master of Science}
\setcounter{tocdepth}{1}

\newcommand{\bspw}{bspw.\xspace}
\newcommand{\bzgl}{bzgl.\xspace}
\newcommand{\bzw}{bzw.\xspace}
\newcommand{\ca}{ca.\xspace}
\newcommand{\gdw}{gdw.\xspace}
\newcommand{\ggf}{ggf.\xspace}
\newcommand{\iA}{i.\,A.\xspace}
\newcommand{\mglw}{mglw.\xspace}
\newcommand{\vs}{vs.\xspace}
\newcommand{\zB}{z.\,B.\xspace}
\renewcommand{\dh}{d.\,h.\xspace}
\newcommand{\ggT}{\ensuremath{\textrm{ggT}}}

\newcommand{\lstbasepath}{unset}

\begin{document}
	\selectlanguage{USenglish}
	\maketitle
	\tableofcontents

	% 1. Semester: Opt
	\cleardoublepage
	\selectlanguage{USenglish}
	\part{Optimization of Static and Dynamic Systems}
	\graphicspath{{./cs/cs/elective/ce/opt/}}
	\renewcommand{\lstbasepath}{./cs/elective/ce/opt}
	\include{cs/elective/ce/opt/content}

	% 2. Semester: RoLe
	\cleardoublepage
	\selectlanguage{USenglish}
	\part{Robot Learning}
	\graphicspath{{./cs/elective/ce/role/}}
	\renewcommand{\lstbasepath}{./cs/elective/ce/role}
	\include{cs/elective/ce/role/content}

	% 3. Semester: DMML, DLAM
	\cleardoublepage
	\selectlanguage{ngerman}
	\part{Data Mining und Maschinelles Lernen}
	\graphicspath{{./cs/elective/iws/dmml/}}
	\renewcommand{\lstbasepath}{./cs/elective/iws/dmml}
	\include{cs/elective/iws/dmml/content}

	\cleardoublepage
	\selectlanguage{USenglish}
	\part{Deep Learning: Architectures and Methods}
	\graphicspath{{./cs/elective/iws/dlam/}}
	\renewcommand{\lstbasepath}{./cs/elective/iws/dlam}
	\include{cs/elective/iws/dlam/content}

	% 4. Semester: PGM, MPC, IVVA
	\cleardoublepage
	\selectlanguage{USenglish}
	\part{Probabilistic Graphical Models}
	\graphicspath{{./cs/elective/vc/pgm/}}
	\renewcommand{\lstbasepath}{./cs/elective/vc/pgm}
	\include{cs/elective/vc/pgm/content}

	\cleardoublepage
	\selectlanguage{USenglish}
	\part{Model-Predictive Control and Machine Learning}
	\graphicspath{{./other/eeit/ccps/mpc/}}
	\renewcommand{\lstbasepath}{./other/eeit/ccps/mpc}
	\include{other/eeit/ccps/mpc/content}

	\cleardoublepage
	\selectlanguage{ngerman}
	\part{Informationsvisualisierung und Visual Analytics}
	\graphicspath{{./cs/elective/vc/ivva/}}
	\renewcommand{\lstbasepath}{./cs/elective/vc/ivva}
	\include{cs/elective/vc/ivva/content}

	% 5. Semester: QC, RL
	\cleardoublepage
	\selectlanguage{USenglish}
	\part{Quantum Computing}
	\graphicspath{{./cs/elective/sf/qc/}}
	\renewcommand{\lstbasepath}{./cs/elective/sf/qc}
	\include{cs/elective/sf/qc/content}

	\cleardoublepage
	\selectlanguage{USenglish}
	\part{Reinforcement Learning}
	\graphicspath{{./cs/elective/ce/rl/}}
	\renewcommand{\lstbasepath}{./cs/elective/ce/rl}
	\include{cs/elective/ce/rl/content}
\end{document}
